Unrecognized States: A Theory of Self-Determination and Foreign Influence

### 2016-06-01 draft.tex
rearrange p/c and s/g payoffs to get them closer together

The payoff functions and all parameters, including probabilities in the war lottery, are common knowledge for all players, and actions are immediately observed by all players at the time they occur. There are an infinite number of discrete periods $t=1,2,\ldots$, with play proceeding until an absorbing state is reached.

Within each period, there is first an investment stage in which first the patron and then the international community each make decisions about how much to invest in altering the payoffs of the `inside actors'---the secessionists and the home government. Then, the inside actors play a ``conflict stage game'' in which they simultaneously choose whether to cede the issue of status, fight, or maintain the status quo.

The payoffs at the end of a period are determined by these actions and six state variables, which keep track of the evolving value of the status quo ($Q_s^t$ and $Q_g^t$), losing the issue of recognition ($L_s^t$ and $L_g^t$) and winning the issue of recognition ($W_s^t$ and $W_g^t$) to the secessionists and government respectively. The game continues into the next period if the inside players selected (status quo, status quo) or (cede, cede). Otherwise, the game ends in an absorbing state as described below.

The status quo payoffs for the secessionists are special in that the transition of this state variable from one period to the next involves an automatic reduction by $\mu$. The steady reduction by $\mu$ represents the costs of non-recognition discussed in the introduction. As the economy in the secessionist region deteriorates, so does the standard of living for the secessionists. Therefore, the payoffs to the secessionists from the status quo evolve according to the law of motion $Q_s^{t} = Q_s^{t-1} - \mu + R_p^t - R_c^t$ where $R_p^t$ is the investment of the patron and $R_c^t$ is the investment of the international community in period $t$.\footnote{Here we anticipate a bit the incentives of the patron and international community to sign the impact of contributions: the patron will want to increase the value of the status quo to the secessionists, while the international community could only desire to reduce this value.}

The other payoffs for the inside actors are also modeled as state variables. However, they remain unchanged from period to period unless players $p$ or $c$ make an investment, to be precise $L_s^t = L_s^{t-1} - R_p^t + R_c^t$ where $L_s^t$ is the per-period payoff to the secessionists from ceding the issue of status. The full set of payoffs for the inside actors are described in Section~\ref{sec:stage}.

Both because of the presence of absorbing states and state variables, this is a dynamic game rather than a standard repeated game.

In each period $t$, play proceeds as follows:

@@

\item $s$ and $g$ play a stage game in which each chooses simultaneously from the following actions: $\{\text{Fight, Status Quo, Cede}\}$. 
\end{enumerate}

In order to fully describe the structure of the game, we first introduce the stage game payoffs.

@@

\noindent Stage game payoffs for players $s$ and $g$ in period $n$, gross of investments by players $p$ and $c$, are:
\vspace{2pt}
\begin{center}
\begin{tabular}{|  c| c | c | c |} 
\hline
  $g\downarrow$,     $ s\rightarrow$  & Fight & Status Quo &Cede \\ \hline
	Fight & $p_g W_g^t + p_s L_g^t - \zeta_g, p_s W_s^t + p_g L_s^t - \zeta_s$ &$p_g W_g^t + p_s L_g^t - \zeta_g, p_s W_s^t + p_g L_s^t - \zeta_s$&$W_g^t, L_s^t$ \\ \hline
	Status Quo& $p_g W_g^t + p_s L_g^t - \zeta_g, p_s W_s^t + p_g L_s^t - \zeta_s$& $\bm{Q_g^t, Q_s^t}$ &$W_g^t, L_s^t$ \\ \hline
	Cede&$L_g^t, W_s^t$& $L_g^t, W_s^t$ &$\bm{Q_g^t, Q_s^t}$ \\ \hline
 \multicolumn{4}{c} {Figure 2: Stage Game Payoffs}\\ 
\end{tabular}
\end{center}
\vspace{2.5pt}

@@

At the start of each period, the patron, $p$, and player $c$ may invest resources to affect the payoffs of $s$ and $g$ in the stage game. For example, if $p$ expends resources $R_p^t$ to increase the government's payoffs from ceding, the previous period's payoff $L_g^{t-1}$ is augmented so that $L_g^t = L_g^{t-1} + R_p^t - R_c^t$. We can write analogous expressions for any of the state variables, recalling that the international community can also make investments to increase whichever of these payoffs they like. $B_p^t$ denotes the total that the patron can devote to interceding in the conflict; we assume that player $c$, which is often a large coalition of states, does not have a binding budget constraint.

The relative strengths of the patron and player $c$'s preferences determine their willingness to spend resources. As we will show, unrecognized statehood emerges as an unhappy but stable middle ground in which all players avoid their least preferred outcomes.

>>>>

The payoff functions and all parameters, including probabilities in the war lottery, are common knowledge for all players, and actions are immediately observed by all players at the time they occur. There are an infinite number of discrete periods $t=1,2,\ldots$. Play proceeds in each period $t$ as follows until an absorbing state is reached.

@@

\item Conflict Stage Game: $s$ and $g$ play a stage game in which each chooses simultaneously from the following actions: $\{\text{Fight, Status Quo, Cede}\}$. 
\end{enumerate}

The payoffs at the end of a period are determined by these actions and six state variables. These state variables keep track of the value of the status quo ($Q_s^t$ and $Q_g^t$), losing ($L_s^t$ and $L_g^t$) and winning the issue of status ($W_s^t$ and $W_g^t$) for the secessionists and government respectively. These values evolve as the patron and player $c$ make investments and the status quo payoffs degrade over time.

The game continues into the next period if the `inside players'---the secessionists and government---selected (status quo, status quo) or (cede, cede). Otherwise, the game ends in an absorbing state as described in Section~\ref{sec:stage}. Both because of the presence of absorbing states and state variables, this is a dynamic game rather than a standard repeated game.

In order to fully describe the structure of the game, we next introduce the payoffs.

At the start of each period, the patron, $p$, and player $c$ may invest resources to affect the payoffs of $s$ and $g$ in the stage game according to Equation~\ref{eqn:motion} below. $B_p^t$ denotes the total that the patron can devote to interceding in the conflict; we assume that player $c$, which is often a large coalition of states, does not have a binding budget constraint.

%The relative strengths of the patron and player $c$'s preferences determine their willingness to spend resources. As we will show, unrecognized statehood emerges as an unhappy but stable middle ground in which all players avoid their least preferred outcomes.

@@

\noindent Stage game payoffs for players $s$ and $g$ in period $t$, gross of investments by players $p$ and $c$, are:
\vspace{2pt}
\begin{center}
\begin{tabular}{|  c| c | c | c |} 
\hline
  $g\downarrow$,     $ s\rightarrow$  & Fight & Status Quo &Cede \\ \hline
	Fight & $p_g W_g^t + p_s L_g^t - \zeta_g, p_s W_s^t + p_g L_s^t - \zeta_s$ &$p_g W_g^t + p_s L_g^t - \zeta_g, p_s W_s^t + p_g L_s^t - \zeta_s$&$W_g^t, L_s^t$ \\ \hline
	Status Quo& $p_g W_g^t + p_s L_g^t - \zeta_g, p_s W_s^t + p_g L_s^t - \zeta_s$& $\bm{Q_g^t, Q_s^t}$ &$W_g^t, L_s^t$ \\ \hline
	Cede&$L_g^t, W_s^t$& $L_g^t, W_s^t$ &$\bm{Q_g^t, Q_s^t}$ \\ \hline
 \multicolumn{4}{c} {Figure 2: Stage Game Payoffs}\\ 
\end{tabular}
\end{center}
\vspace{2.5pt}

All the state variables except for the secessionists' status quo payoffs remain unchanged from period to period unless players $p$ or $c$ make an investment. To be precise, we have\footnote{Here we anticipate a bit the incentives of the patron and international community to sign the impact of contributions in each case.}
	\begin{equation}
	\begin{aligned}
		L_s^t = L_s^{t-1} - R_p^t + R_c^t \hskip.5in L_g^t = L_g^{t-1} + R_p^t - R_c^t \\
		W_s^t = W_s^{t-1} + R_p^t - R_c^t \hskip.5in W_g^t = W_g^{t-1} - R_p^t + R_c^t \\
		Q_g^t = Q_g^{t-1} - R_p^t + R_c^t \hskip.5in Q_s^t = Q_s^{t-1} - \mu + R_p^t - R_c^t
		\label{eqn:motion}
	\end{aligned}
	\end{equation}

The status quo payoffs for the secessionists are special in that the transition of this state variable from one period to the next involves an automatic reduction by $\mu$. The steady reduction by $\mu$ represents the costs of non-recognition discussed in the introduction. As the economy in the secessionist region deteriorates, so does the standard of living for the secessionists.

@@
